---
title: "multi_state_quarto"
format: docx
editor: visual
---
```{r, setup, include=FALSE}
# hide messages/warnings and set default figure size & resolution
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  # size in inches (use sensible values for Word)
  fig.width = 9,
  fig.height = 6,
  dpi = 300,
  # for Word: control rendered image size (inches); change if you prefer
  out.width = "6in"
)

```

## HERON-UK-02-002-CVDValveReplacement

## Import Results

```{r, echo = FALSE}
library(ggplot2)
results <- omopgenerics::importSummarisedResult(here::here("rawData"))
```

## Incidence

```{r, echo = FALSE}
data_inc <- results |>
  omopgenerics::filterSettings(result_type == "incidence")
```
Age-standardised incidence by socioeconomic status and sex

```{r, echo = FALSE}
data_inc |>
  omopgenerics::filterSettings(standardised_incidence == "TRUE") |>
  omopgenerics::filterAdditional(analysis_interval == "overall") |>
  #omopgenerics::filterSettings(denominator_sex == "Both") |>
  omopgenerics::filterStrata(ses != "Missing") |>
  IncidencePrevalence::plotIncidence(x = "ses", colour = "denominator_sex", facet = cdm_name~outcome_cohort_name) +
  facet_grid(cdm_name ~ outcome_cohort_name, scales = "free", space = "free") +
  coord_cartesian(ylim = c(20, 160)) 



```


Incidence of Aortic Stenosis by age group and socioeconomic status
```{r, echo = FALSE}
data_inc |>
  omopgenerics::filterSettings(standardised_incidence == "FALSE") |>
  omopgenerics::filterAdditional(analysis_interval != "overall") |>
  omopgenerics::filterSettings(denominator_sex == "Both") |>
  omopgenerics::filterStrata(ses != "Missing") |>
  omopgenerics::filterGroup(outcome_cohort_name == "aortic_stenosis") |>
  IncidencePrevalence::plotIncidence(colour = "denominator_age_group", facet = ses~cdm_name, line = TRUE, ribbon = TRUE) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
  
```
Incidence of Aortic Valve Replacement by age group and socioeconomic status

```{r, echo = FALSE}
data_inc |>
  omopgenerics::filterSettings(standardised_incidence == "FALSE") |>
  omopgenerics::filterAdditional(analysis_interval != "overall") |>
  omopgenerics::filterSettings(denominator_sex == "Both") |>
  omopgenerics::filterStrata(ses != "Missing") |>
  omopgenerics::filterGroup(outcome_cohort_name == "aortic_valve_replacement") |>
  IncidencePrevalence::plotIncidence(colour = "denominator_age_group", facet = ses~cdm_name, line = TRUE, ribbon = TRUE) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```
## Characteristics 

```{r, echo = FALSE}
data_char <- results |>
  omopgenerics::filterSettings(result_type == "summarise_characteristics")
```
### Characteristics of healthy population
```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "healthy") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record"))) |>
  CohortCharacteristics::tableCharacteristics()

```
### Characteristics of aortic stenosis patients
```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "healthy_to_as") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort"  ))) |>
  CohortCharacteristics::tableCharacteristics()

```
### Characteristics of patients dying after AVR 
```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "healthy_to_death") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort"))) |>
  CohortCharacteristics::tableCharacteristics()

```
### Characteristics of patients receiving AVR after AS diagnosis
```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "as_to_avr") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort"))) |>
  CohortCharacteristics::tableCharacteristics()

```
### Characteristics of patients dying after AS diagnosis without receiving AVR
```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "as_to_death") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort"))) |>
  CohortCharacteristics::tableCharacteristics()

```
### Characteristics of patients dying after AVR 
```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "avr_to_death") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort"))) |>
  CohortCharacteristics::tableCharacteristics()

```

### age and sex distribution 
```{r}
data_char |>
  omopgenerics::tidy() |> 
  dplyr::filter(variable_name %in% c("Age group", "Sex") ) |> 
  dplyr::select(cdm_name, cohort_name, variable_name, variable_level, count) 
  
    
```




## Multi state model
```{r, echo = FALSE}
data_hr <- results |>
  omopgenerics::filterSettings(result_type == "hr_summary")
```
### only age model
```{r, echo = FALSE}
data_hr |> omopgenerics::filterStrata(.data$ref_age == "70") |>
      omopgenerics::tidy()|>
      dplyr::mutate(rel_age = as.numeric(.data$variable_level), 
                    transition = dplyr::case_when(
                      .data$transition == "healthy_to_as" ~ "Healthy to AS",
                      .data$transition == "healthy_to_death" ~ "Healthy to Death",
                      .data$transition == "as_to_avr" ~ "AS to AVR",
                      .data$transition == "as_to_death" ~ "AS to Death",
                      .data$transition == "avr_to_death" ~ "AVR to Death"
                      
                    )) |>
      
      visOmopResults::scatterPlot(x = "rel_age", 
                                  y = "hazard_ratio", 
                                  line = TRUE, 
                                  point = FALSE, 
                                  ribbon = FALSE,
                                  facet =  transition~cdm_name,
                                  colour = "cdm_name") +
      geom_line(aes(y = lower_hr), linetype = "dashed", linewidth = 0.6) +
      geom_line(aes(y = upper_hr), linetype = "dashed", linewidth = 0.6) +
      coord_cartesian(ylim = c(0, 10)) +
  xlab("Relative age") +
      labs(
        title = "Relative hazard ratios: age only model",
        subtitle = "Reference age: 70"
      )
```
### Other models
```{r, echo = FALSE}
data_hr |> 
  omopgenerics::filterStrata(.data$ref_age != "70") |>
  dplyr::filter(variable_name != "age") |>
  omopgenerics::tidy() |>
  dplyr::mutate(transition = dplyr::case_when(
                      .data$transition == "healthy_to_as" ~ "Healthy to AS",
                      .data$transition == "healthy_to_death" ~ "Healthy to Death",
                      .data$transition == "as_to_avr" ~ "AS to AVR",
                      .data$transition == "as_to_death" ~ "AS to Death",
                      .data$transition == "avr_to_death" ~ "AVR to Death"
                      
                    )) |>
  visOmopResults::scatterPlot(x = "variable_level", 
                              y = "hazard_ratio", 
                              line = FALSE,
                              point = TRUE, 
                              ribbon = FALSE,
                              facet = transition~cdm_name, 
                              colour = "model_name",
                              ymin = "lower_hr", 
                              ymax = "upper_hr") +
  geom_hline(yintercept = 1, linetype = "dashed", linewidth = 0.6, color = "black") +
  coord_cartesian(ylim = c(0,3)) +
  xlab("Variable") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
      
```
```{r, echo = FALSE}
estimates <- c("Hazard Ratio [95% CI]" = "<hazard_ratio> [<lower_hr>, <upper_hr>]", 
               "P Value" = "<p_value>", 
               "Standard Error" = "<se_coef>"
               ) 
data_hr |> 
  omopgenerics::filterStrata(.data$ref_age != "70") |>
  dplyr::filter(variable_name != "age") |>
  visOmopResults::visOmopTable(header = "cdm_name", groupColumn = "transition", hide = "ref_age",
                               columnOrder = c("model_name", "variable_name", "variable_level",
                                               "estimate_name", "estimate_value"),
                               estimateName = estimates, 
                               .options = list(decimals = c("numeric" = 3))
                               )
```

