---
title: "multi_state_quarto"
format: docx
editor: visual
---

```{r, setup, include=FALSE}
# hide messages/warnings and set default figure size & resolution
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  # size in inches (use sensible values for Word)
  fig.width = 9,
  fig.height = 6,
  dpi = 300,
  # for Word: control rendered image size (inches); change if you prefer
  out.width = "6in"
)

```

## HERON-UK-02-002-CVDValveReplacement

## Import Results

```{r, echo = FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
results <- omopgenerics::importSummarisedResult(here::here("rawData"))


renameCdm <- function(res){
  res |>
    mutate(cdm_name =
             case_when(str_detect(tolower(cdm_name), "aurum|heron_cdm")  ~ "CPRD Aurum",
                       str_detect(tolower(cdm_name), "dataloch") ~ "DataLoch Lothian",
                       .default = cdm_name
             )
    )
}
results <- results |>
  renameCdm()

```

## Incidence

```{r, echo = FALSE}
data_inc <- results |>
  omopgenerics::filterSettings(result_type == "incidence")
data_inc <- data_inc |> 
  mutate(group_level = str_replace(
    group_level, 
    "aortic_stenosis", "Aortic stenosis")) |> 
  mutate(group_level = str_replace(
    group_level, 
    "aortic_valve_replacement", "Aortic valve replacement"))
```

### Age-standardised incidence by socioeconomic status

```{r, echo = FALSE}
sesStandarisedPlot <- function() {
  working_plot <- data_inc |>
    omopgenerics::filterSettings(standardised_incidence == "TRUE") |>
    omopgenerics::filterAdditional(analysis_interval == "overall") |>
    omopgenerics::filterSettings(denominator_sex == "Both") |>
    omopgenerics::filterStrata(ses != "Missing") |>
    dplyr::filter(strata_level != "overall") |> 
    IncidencePrevalence::plotIncidence(colour = "ses", 
                                       x = "ses",
                                       facet = outcome_cohort_name~cdm_name, 
                                       line = TRUE, ribbon = TRUE) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 0, vjust = 0.5),
        strip.background = element_rect(
          fill = "gray94"
    )) +
    facet_grid(outcome_cohort_name~cdm_name, scales = "free_y")+
    labs(colour = "Socioeconomic status",
         fill = "Socioeconomic status") +
    theme(legend.position = "top")+
    xlab("Socioeconomic status") 
  
  working_plot@data <- working_plot@data |> 
    dplyr::mutate(ses = dplyr::if_else(
      ses == "1",
      "1 (least deprived)", ses)) |> 
    dplyr::mutate(ses = dplyr::if_else(
      ses == "5",
      "5 (most deprived)", ses))
  
  working_plot
  
  working_plot
}
sesStandarisedPlot()
```

### Incidence by age and socioeconomic status

```{r, echo = FALSE}
sesIncPlotAge <- function() {
  working_plot <- data_inc |>
    omopgenerics::filterSettings(standardised_incidence == "FALSE") |>
    omopgenerics::filterAdditional(analysis_interval == "overall") |>
    omopgenerics::filterSettings(denominator_sex == "Both") |>
    omopgenerics::filterSettings(denominator_age_group != "20 to 150") |>
    omopgenerics::filterStrata(ses != "Missing") |>
    dplyr::filter(strata_level != "overall") |> 
    IncidencePrevalence::plotIncidence(colour = "ses", 
                                       x = "denominator_age_group",
                                       facet = ses~cdm_name, 
                                       line = TRUE, ribbon = TRUE) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        strip.background = element_rect(
          fill = "gray94"
    )) +
    facet_grid(outcome_cohort_name~cdm_name, scales = "free_y")+
    labs(colour = "Socioeconomic status",
         fill = "Socioeconomic status") +
    theme(legend.position = "top")+
    xlab("Age") 
  
  working_plot@data <- working_plot@data |> 
    dplyr::mutate(ses = dplyr::if_else(
      ses == "1",
      "1 (least deprived)", ses)) |> 
    dplyr::mutate(ses = dplyr::if_else(
      ses == "5",
      "5 (most deprived)", ses))
  
  working_plot
  
  working_plot
}
sesIncPlotAge()
```

### Incidence by year and socioeconomic status

```{r, echo = FALSE}
sesIncPlot <- function(outcome) {
working_plot <- data_inc |>
  omopgenerics::filterSettings(standardised_incidence == "FALSE") |>
  omopgenerics::filterAdditional(analysis_interval != "overall") |>
  omopgenerics::filterSettings(denominator_sex == "Both") |>
  omopgenerics::filterStrata(ses != "Missing") |>
  dplyr::filter(strata_level != "overall") |> 
  omopgenerics::filterSettings(denominator_age_group == "20 to 150") |> 
  IncidencePrevalence::plotIncidence(colour = "ses", 
                                     facet = denominator_age_group~cdm_name, 
                                     line = TRUE, ribbon = TRUE) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        strip.background = element_rect(
          fill = "gray94"
    ))+
  facet_grid(outcome_cohort_name~cdm_name, scales = "free_y") +
  ylim(c(0, NA)) + 
  labs(colour = "Socioeconomic status",
       fill = "Socioeconomic status") +
  theme(legend.position = "top")

working_plot@data <- working_plot@data |> 
  dplyr::mutate(denominator_age_group = dplyr::if_else(
    denominator_age_group == "20 to 150",
    "Overall", denominator_age_group))|> 
  dplyr::mutate(ses = dplyr::if_else(
    ses == "1",
    "1 (least deprived)", ses)) |> 
  dplyr::mutate(ses = dplyr::if_else(
    ses == "5",
    "5 (most deprived)", ses))


working_plot
}
sesIncPlot()
```

Age-standardised incidence by socioeconomic status and sex

```{r, echo = FALSE}
data_inc |>
  omopgenerics::filterSettings(standardised_incidence == "TRUE") |>
  omopgenerics::filterAdditional(analysis_interval == "overall") |>
  #omopgenerics::filterSettings(denominator_sex == "Both") |>
  omopgenerics::filterStrata(ses != "Missing") |>
  IncidencePrevalence::plotIncidence(x = "ses", colour = "denominator_sex", facet = outcome_cohort_name ~ cdm_name) +
  facet_wrap(outcome_cohort_name ~ cdm_name, scales = "free_y") 

```

Incidence of Aortic Stenosis by age group and socioeconomic status

```{r, echo = FALSE}
data_inc |>
  omopgenerics::filterSettings(standardised_incidence == "FALSE") |>
  omopgenerics::filterAdditional(analysis_interval != "overall") |>
  omopgenerics::filterSettings(denominator_sex == "Both") |>
  omopgenerics::filterStrata(ses != "Missing") |>
  omopgenerics::filterGroup(outcome_cohort_name == "Aortic stenosis") |>
  IncidencePrevalence::plotIncidence(colour = "denominator_age_group", facet = ses~cdm_name, line = TRUE, ribbon = TRUE) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
  
```

Incidence of Aortic Valve Replacement by age group and socioeconomic status

```{r, echo = FALSE}
data_inc |>
  omopgenerics::filterSettings(standardised_incidence == "FALSE") |>
  omopgenerics::filterAdditional(analysis_interval != "overall") |>
  omopgenerics::filterSettings(denominator_sex == "Both") |>
  omopgenerics::filterStrata(ses != "Missing") |>
  omopgenerics::filterGroup(outcome_cohort_name == "Aortic valve replacement") |>
  IncidencePrevalence::plotIncidence(colour = "denominator_age_group", facet = ses~cdm_name, line = TRUE, ribbon = TRUE) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```

## Characteristics

```{r, echo = FALSE}
data_char <- results |>
  omopgenerics::filterSettings(result_type == "summarise_characteristics")
```

### Characteristics of healthy population

```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "healthy") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record", "Ses", "Cohort end date","Future observation", "Days in cohort", "Ses", "Ethnicity" ))) |>
  CohortCharacteristics::tableCharacteristics(header = c("cdm_name", "cohort_name", "ses"))

```

### Characteristics of aortic stenosis patients

```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "healthy_to_as") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort", "Ses", "Ethnicity"  ))) |>
  CohortCharacteristics::tableCharacteristics(header = c("cdm_name", "cohort_name", "ses"))

```

### Characteristics of patients dying after AVR

```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "healthy_to_death") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort", "Ses", "Ethnicity"))) |>
  
  CohortCharacteristics::tableCharacteristics(header = c("cdm_name", "cohort_name", "ses"))

```

### Characteristics of patients receiving AVR after AS diagnosis

```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "as_to_avr") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort", "Ses", "Ethnicity"))) |>
  CohortCharacteristics::tableCharacteristics(header = c("cdm_name", "cohort_name", "ses"))

```

### Characteristics of patients dying after AS diagnosis without receiving AVR

```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "as_to_death") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort", "Ses", "Ethnicity"))) |>
  CohortCharacteristics::tableCharacteristics(header = c("cdm_name", "cohort_name", "ses"))

```

### Characteristics of patients dying after AVR

```{r, echo = FALSE}
data_char |>
  omopgenerics::filterGroup(cohort_name == "avr_to_death") |>
  dplyr::filter(!(variable_name %in% c("Number records", "Days to next record","Cohort end date","Future observation", "Days in cohort","Ses", "Ethncity"))) |>
  CohortCharacteristics::tableCharacteristics(header = c("cdm_name", "cohort_name", "ses"))

```

## Multi state model

```{r, echo = FALSE}
data_hr <- results |>
  omopgenerics::filterSettings(result_type == "hr_summary") |>
  omopgenerics::tidy()|>
  dplyr::mutate(transition = dplyr::case_when(
                      .data$transition == "healthy_to_as" ~ "General Pop to AS",
                      .data$transition == "healthy_to_death" ~ "General Pop to Death",
                      .data$transition == "as_to_avr" ~ "AS to AVR",
                      .data$transition == "as_to_death" ~ "AS to Death",
                      .data$transition == "avr_to_death" ~ "AVR to Death"
                      
                    )) |>
    dplyr::mutate(age_limit = dplyr::case_when(
                      age_limit == "overall" ~ "Overall", 
                      age_limit == "below_80" ~ "Below 80", 
                      age_limit == "over_80" ~ "Over 80"
                      
                      
                    ), 
                  transition = factor(transition, levels = c(
      "General Pop", "General Pop to AS", "General Pop to Death",
      "AS to AVR", "AS to Death", "AVR to Death"
    ))
  ) 
```

### only age model

```{r, echo = FALSE}
data_hr |> dplyr::filter(.data$ref_age == "70") |>
      dplyr::mutate(rel_age = as.numeric(.data$variable_level) )|> 
      visOmopResults::scatterPlot(x = "rel_age", 
                                  y = "hazard_ratio", 
                                  line = TRUE, 
                                  point = FALSE, 
                                  ribbon = FALSE,
                                  facet =  transition~age_limit,
                                  colour = "cdm_name") +
      geom_line(aes(y = lower_hr), linetype = "dashed", linewidth = 0.6) +
      geom_line(aes(y = upper_hr), linetype = "dashed", linewidth = 0.6) +
  xlab("Relative age") +
      labs(
        title = "Relative hazard ratios: age only model",
        subtitle = "Reference age: 70"
      ) +
  facet_grid(transition~age_limit, scales = "free_y")
```
```{r}

data_hr |> dplyr::filter(.data$ref_age != "70", variable_name != "age",model_name == "age_sex" ) |>
  visOmopResults::scatterPlot(x = "transition", 
                              y = "hazard_ratio", 
                              line = FALSE,
                              point = TRUE, 
                              ribbon = FALSE,
                              facet = age_limit~cdm_name, 
                              colour = "variable_level",
                              ymin = "lower_hr", 
                              ymax = "upper_hr") +
  facet_grid(age_limit~cdm_name, scales = "free_y") +
  geom_hline(yintercept = 1, linetype = "dashed", linewidth = 0.6, color = "black") +
  scale_colour_manual(values = c( "darkblue"))+
  labs(title = "Hazard Ratios: Males vs Females (Adjusted for age)") +
  theme(legend.position = "none")


```

### ses models

```{r}
data_hr2 <- data_hr |> 
  dplyr::filter(.data$ref_age != "70") |>
  dplyr::filter(variable_name != "age") 
```

```{r, echo = FALSE}
data_hr2 |> 

  dplyr::filter(model_name == "ses_age_sex", 
                !(variable_level %in% c("Male", "Missing")) ) |>
  
  visOmopResults::scatterPlot(x = "variable_level", 
                              y = "hazard_ratio", 
                              line = FALSE,
                              point = TRUE, 
                              ribbon = FALSE,
                              facet = transition~age_limit, 
                              colour = "variable_level",
                              ymin = "lower_hr", 
                              ymax = "upper_hr") +
  geom_hline(yintercept = 1, linetype = "dashed", linewidth = 0.6, color = "black") +
  xlab("Socioeconomic status") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )+
facet_grid(transition~age_limit, scales = "free_y")+
scale_colour_discrete(
name = "Socioeconomic status",
labels = c(
"1" = "1 (Least deprived)",
"2" = "2",
"3" = "3",
"4" = "4"
)
) +
scale_fill_discrete(
name = "Socioeconomic status",
labels = c(
"1" = "1 (Least deprived)",
"2" = "2",
"3" = "3",
"4" = "4"
)
) +
labs(
title = "Hazard ratios by socioeconomic status, adjusted for age and sex.",
subtitle = "Reference is socioeconomic status 5 (Most deprived)."
)
```

```{r, echo = FALSE}
estimates <- c("Hazard Ratio [95% CI]" = "<hazard_ratio> [<lower_hr>, <upper_hr>]", 
               "P Value" = "<p_value>", 
               "Standard Error" = "<se_coef>"
               ) 
data_hr |> 
  dplyr::filter(.data$ref_age != "70") |>
  dplyr::filter(variable_name != "age", model_name %in% c("ses", "ses_age_sex" )) |>
  tidyr::pivot_longer(cols = c("hazard_ratio","lower_hr", "upper_hr","p_value", "se_coef"),
                      names_to = "estimate_name", 
                      values_to = "estimate_value") |>
  dplyr::mutate(estimate_type = "numeric", 
                variable = paste0(
  stringr::str_to_sentence(.data$variable_name),
  ": ",
  .data$variable_level
), 
model = dplyr::if_else(.data$model_name == "ses", "Unadjusted", "Adjusted for age and sex")) |>
  dplyr::arrange(.data$model_name) |>
  dplyr::select(cdm_name, model, variable, estimate_name, estimate_value, estimate_type, transition, age_limit) |>
  visOmopResults::visTable(header = c("cdm_name", "age_limit"), groupColumn = "transition", hide = c("ref_age","estimate_type"),
                               
                               estimateName = estimates, 
                               .options = list(decimals = c("numeric" = 3))
                               )
```
# Comorbidities

```{r}
grid_lines <- data_char |>
  dplyr::filter(variable_name %in% c("Comorbidities", "Ckd stage")) |>
  omopgenerics::tidy() |>
  dplyr::filter(ses == "overall") |>
  dplyr::mutate(
    transition = dplyr::case_when(
      .data$cohort_name == "healthy"          ~ "General Pop",
      .data$cohort_name == "healthy_to_as"    ~ "General Pop to AS",
      .data$cohort_name == "healthy_to_death" ~ "General Pop to Death",
      .data$cohort_name == "as_to_avr"        ~ "AS to AVR",
      .data$cohort_name == "as_to_death"      ~ "AS to Death",
      .data$cohort_name == "avr_to_death"     ~ "AVR to Death"
    ),
    transition = factor(transition, levels = c(
      "General Pop", "General Pop to AS", "General Pop to Death",
      "AS to AVR", "AS to Death", "AVR to Death"
    ))
  ) |>
  dplyr::distinct(transition) |>
  tidyr::crossing(yintercept = c(20, 40, 60, 80)) |>
  dplyr::mutate(label = paste0(yintercept, "%"))

data_char |>
  dplyr::filter(variable_name %in% c("Comorbidities", "Ckd stage") & .data$variable_level != "Missing" & !(.data$variable_level %in% c("Stage 1", "Stage 2", "Stage 3"))) |>
  dplyr::mutate(variable_level = dplyr::if_else(.data$variable_name == "Ckd stage", paste0("CKD ", .data$variable_level), .data$variable_level)) |>
  omopgenerics::tidy() |>
  dplyr::filter(ses == "overall") |>
  dplyr::mutate(
    transition = dplyr::case_when(
      .data$cohort_name == "healthy"          ~ "General Pop",
      .data$cohort_name == "healthy_to_as"    ~ "General Pop to AS",
      .data$cohort_name == "healthy_to_death" ~ "General Pop to Death",
      .data$cohort_name == "as_to_avr"        ~ "AS to AVR",
      .data$cohort_name == "as_to_death"      ~ "AS to Death",
      .data$cohort_name == "avr_to_death"     ~ "AVR to Death"
    ),
    transition = factor(transition, levels = c(
      "General Pop", "General Pop to AS", "General Pop to Death",
      "AS to AVR", "AS to Death", "AVR to Death"
    ))
  ) |>
  ggplot2::ggplot(ggplot2::aes(x = variable_level, y = percentage, fill = variable_level)) +
  ggplot2::geom_col(width = 1, colour = "white", linewidth = 0.3) +
  ggplot2::geom_hline(
    data = grid_lines,
    ggplot2::aes(yintercept = yintercept, group = transition),
    colour = "grey60", linewidth = 0.3, linetype = "dashed", inherit.aes = FALSE
  ) +
  
  ggplot2::geom_text(
  data = dplyr::filter(grid_lines, yintercept != 20),
  ggplot2::aes(x = 2, y = yintercept, label = label),
  colour = "grey40", size = 2.5, inherit.aes = FALSE
)+
  ggplot2::coord_polar() +
  ggplot2::scale_y_continuous(limits = c(0, 100)) +
  ggplot2::facet_wrap(~ transition) +
  ggplot2::scale_fill_viridis_d(option = "turbo") +
  ggplot2::labs(
    x = NULL, y = NULL, fill = "Comorbidity",
    title = "Comorbidities"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.text.x     = ggplot2::element_blank(),
    axis.text.y     = ggplot2::element_blank(),
    panel.grid      = ggplot2::element_blank(),
    legend.position = "right"
  ) 
```

