---
title: ""
format:
  html:
    page-layout: full
    css: styles.css
editor: visual
fig-width: 12  
---

::: {style="display: flex; justify-content: space-between; align-items: center; margin-bottom: 2em; margin-top: 1em;"}
::: {style="flex-grow: 1;"}
# Congenital Aortic Stenosis
:::
::: {style="display: flex; gap: 1em; align-items: center;"}
![](hds_logo.svg){width="80px"} ![](ohdsi_logo.svg){width="80px"}
:::
:::


```{r loading results, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(gt)
library(ggplot2)

# Load all CSV results
all_results <- list.files("../results", pattern = "\\.csv$", full.names = TRUE) |>
  lapply(read.csv) |>
  bind_rows()

# Helper function to process survival curves with dynamic column renaming
process_survival_curves <- function(data, strata_filter = "overall") {
  raw_data <- data |>
    filter(
      result_id %in% c(12, 16),
      variable_name == "outcome",
      grepl("estimate", estimate_name, ignore.case = TRUE),
      if(strata_filter == "overall") strata_name == "overall" else grepl(strata_filter, strata_name)
    ) |>
    select(cdm_name, group_level, variable_level, any_of("strata_level"), 
           estimate_name, estimate_value, additional_level) |>
    mutate(time = as.numeric(additional_level),
           estimate_value = as.numeric(estimate_value)) |>
    pivot_wider(names_from = estimate_name, values_from = estimate_value, values_fn = first)
  
  # Return empty dataframe if no data
  if(nrow(raw_data) == 0) {
    return(data.frame())
  }
  
  # Dynamic column renaming - build named vector for rename
  cols <- names(raw_data)
  result <- raw_data |> rename(target_cohort = group_level, outcome = variable_level)
  
  # Rename survival probability column
  if("estimate" %in% cols) {
    result <- result |> rename(survival_prob = estimate)
  } else {
    est_match <- cols[grepl("^estimate$|survival.*estimate", cols, ignore.case = TRUE)]
    if(length(est_match) > 0 && !is.na(est_match[1])) {
      result <- result |> rename(survival_prob = !!est_match[1])
    }
  }
  
  # Rename CI lower column
  if("estimate_95CI_lower" %in% cols) {
    result <- result |> rename(ci_lower = estimate_95CI_lower)
  } else {
    ci_match <- cols[grepl("lower|min", cols, ignore.case = TRUE)]
    if(length(ci_match) > 0 && !is.na(ci_match[1])) {
      result <- result |> rename(ci_lower = !!ci_match[1])
    }
  }
  
  # Rename CI upper column
  if("estimate_95CI_upper" %in% cols) {
    result <- result |> rename(ci_upper = estimate_95CI_upper)
  } else {
    ci_match <- cols[grepl("upper|max", cols, ignore.case = TRUE)]
    if(length(ci_match) > 0 && !is.na(ci_match[1])) {
      result <- result |> rename(ci_upper = !!ci_match[1])
    }
  }
  
  result
}

# Helper function to format cohort labels
format_cohort_labels <- function(cohort) {
  case_when(
    cohort == "aortic_insufficiency_congenital" ~ "Aortic Insufficiency Congenital",
    cohort == "aortic_stenosis_congenital" ~ "Aortic Stenosis",
    cohort == "congenital_aortic_stenosis" ~ "Congenital Aortic Stenosis",
    cohort == "congenital_aortic_valve_disease" ~ "Congenital Aortic Valve Disease",
    TRUE ~ gsub("_", " ", cohort) |> tools::toTitleCase()
  )
}

# Prepare survival summary table
survival_table_raw <- all_results |>
  filter(result_id %in% c(14, 18), variable_name == "outcome", strata_name == "overall") |>
  select(cdm_name, group_level, variable_level, estimate_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  pivot_wider(names_from = estimate_name, values_from = estimate_value, values_fn = first)

table_cols <- names(survival_table_raw)
survival_table <- survival_table_raw |>
  mutate(
    median_survival_formatted = if("median_survival" %in% table_cols && 
                                     all(c("median_survival_95CI_lower", "median_survival_95CI_higher") %in% table_cols)) {
      ifelse(is.na(median_survival), "–", 
             sprintf("%.2f (%.2f, %.2f)", median_survival, median_survival_95CI_lower, median_survival_95CI_higher))
    } else "–",
    restricted_mean_survival_formatted = if(all(c("restricted_mean_survival", "restricted_mean_survival_95CI_lower", 
                                                   "restricted_mean_survival_95CI_upper") %in% table_cols)) {
      sprintf("%.2f (%.2f, %.2f)", restricted_mean_survival, restricted_mean_survival_95CI_lower, restricted_mean_survival_95CI_upper)
    } else NA_character_
  ) |>
  select(`CDM name` = cdm_name, `Target cohort` = group_level, `Outcome name` = variable_level,
         any_of(c("Number records" = "number_records_count", "Number events" = "n_events_count")),
         `Median survival (95% CI)` = median_survival_formatted,
         `Restricted mean survival (95% CI)` = restricted_mean_survival_formatted) |>
  arrange(`CDM name`, `Target cohort`, `Outcome name`)

# Prepare survival curves for different stratifications
survival_curves <- process_survival_curves(all_results, "overall")
survival_by_age <- process_survival_curves(all_results, "age_group") |> rename(age_group = strata_level)
survival_by_sex <- process_survival_curves(all_results, "sex") |> rename(sex = strata_level)

# Get total subjects per cohort for percentage calculation
total_subjects <- all_results |>
  filter(result_id == 8, variable_name == "Number subjects", estimate_name == "count") |>
  select(cdm_name, group_level, n_total = estimate_value) |>
  mutate(n_total = as.numeric(n_total))

# Prepare comorbidities (Result ID 9 - pre-index conditions) - Top 10
comorbidities_pre <- all_results |>
  filter(result_id == 9, 
         estimate_name == "count",
         variable_level == "any_time_prior_to_day_prior") |>
  select(cdm_name, group_level, variable_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  left_join(total_subjects, by = c("cdm_name", "group_level")) |>
  mutate(percentage = (estimate_value / n_total) * 100,
         formatted = sprintf("%d (%.1f%%)", estimate_value, percentage)) |>
  group_by(group_level, variable_name) |>
  summarise(
    across(c(cdm_name, formatted), ~paste(., collapse = " | ")),
    total_count = sum(estimate_value),
    .groups = "drop"
  ) |>
  group_by(group_level) |>
  arrange(desc(total_count)) |>
  slice_head(n = 10) |>
  ungroup() |>
  select(`Target Cohort` = group_level, Condition = variable_name, `N (%) by Database` = formatted)

# Prepare index date admissions (Result ID 10)
admissions_index <- all_results |>
  filter(result_id == 10, 
         estimate_name == "count",
         variable_level == "on_index_date") |>
  select(cdm_name, group_level, variable_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  left_join(total_subjects, by = c("cdm_name", "group_level")) |>
  mutate(percentage = (estimate_value / n_total) * 100,
         formatted = sprintf("%d (%.1f%%)", estimate_value, percentage)) |>
  select(group_level, variable_name, cdm_name, formatted) |>
  pivot_wider(names_from = cdm_name, values_from = formatted) |>
  select(`Target Cohort` = group_level, `Admission Type` = variable_name, everything())

# Prepare index date procedures (Result ID 11) - Top 10
procedures_index <- all_results |>
  filter(result_id == 11, 
         estimate_name == "count",
         variable_level == "on_index_date") |>
  select(cdm_name, group_level, variable_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  left_join(total_subjects, by = c("cdm_name", "group_level")) |>
  mutate(percentage = (estimate_value / n_total) * 100,
         formatted = sprintf("%d (%.1f%%)", estimate_value, percentage)) |>
  group_by(group_level, variable_name) |>
  summarise(
    across(c(cdm_name, formatted), ~paste(., collapse = " | ")),
    total_count = sum(estimate_value),
    .groups = "drop"
  ) |>
  group_by(group_level) |>
  arrange(desc(total_count)) |>
  slice_head(n = 10) |>
  ungroup() |>
  select(`Target Cohort` = group_level, Procedure = variable_name, `N (%) by Database` = formatted)

# Prepare index date medications (Result ID 12) - Top 10
medications_index <- all_results |>
  filter(result_id == 12, 
         estimate_name == "count",
         variable_level == "on_index_date") |>
  select(cdm_name, group_level, variable_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  left_join(total_subjects, by = c("cdm_name", "group_level")) |>
  mutate(percentage = (estimate_value / n_total) * 100,
         formatted = sprintf("%d (%.1f%%)", estimate_value, percentage)) |>
  group_by(group_level, variable_name) |>
  summarise(
    across(c(cdm_name, formatted), ~paste(., collapse = " | ")),
    total_count = sum(estimate_value),
    .groups = "drop"
  ) |>
  group_by(group_level) |>
  arrange(desc(total_count)) |>
  slice_head(n = 10) |>
  ungroup() |>
  select(`Target Cohort` = group_level, Medication = variable_name, `N (%) by Database` = formatted)

# Prepare follow-up period events - Top 10
followup_conditions <- all_results |>
  filter(result_id == 9, 
         estimate_name == "count",
         grepl("day_post", variable_level)) |>
  select(cdm_name, group_level, variable_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  left_join(total_subjects, by = c("cdm_name", "group_level")) |>
  mutate(percentage = (estimate_value / n_total) * 100,
         formatted = sprintf("%d (%.1f%%)", estimate_value, percentage)) |>
  group_by(group_level, variable_name) |>
  summarise(
    across(c(cdm_name, formatted), ~paste(., collapse = " | ")),
    total_count = sum(estimate_value),
    .groups = "drop"
  ) |>
  group_by(group_level) |>
  arrange(desc(total_count)) |>
  slice_head(n = 10) |>
  ungroup() |>
  mutate(event_type = "Condition")

followup_procedures <- all_results |>
  filter(result_id == 11, 
         estimate_name == "count",
         grepl("day_post", variable_level)) |>
  select(cdm_name, group_level, variable_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  left_join(total_subjects, by = c("cdm_name", "group_level")) |>
  mutate(percentage = (estimate_value / n_total) * 100,
         formatted = sprintf("%d (%.1f%%)", estimate_value, percentage)) |>
  group_by(group_level, variable_name) |>
  summarise(
    across(c(cdm_name, formatted), ~paste(., collapse = " | ")),
    total_count = sum(estimate_value),
    .groups = "drop"
  ) |>
  group_by(group_level) |>
  arrange(desc(total_count)) |>
  slice_head(n = 10) |>
  ungroup() |>
  mutate(event_type = "Procedure")

# Baseline characteristics summary
baseline_summary <- all_results |>
  filter(result_id == 8, strata_name == "overall",
         variable_name %in% c("Number subjects", "Age", "Sex", "Days in cohort")) |>
  select(cdm_name, group_level, variable_name, variable_level, estimate_name, estimate_value) |>
  mutate(estimate_value = as.numeric(estimate_value)) |>
  pivot_wider(names_from = estimate_name, values_from = estimate_value, values_fn = first) |>
  mutate(
    display_value = case_when(
      variable_name == "Number subjects" ~ as.character(round(count)),
      variable_name == "Age" ~ sprintf("%.1f (%.1f-%.1f)", median, q25, q75),
      variable_name == "Sex" ~ sprintf("%d (%.1f%%)", round(count), percentage),
      variable_name == "Days in cohort" ~ sprintf("%.0f (%.0f-%.0f)", median, q25, q75),
      TRUE ~ as.character(round(count))
    )
  ) |>
  unite(var_display, variable_name, variable_level, sep = ": ", na.rm = TRUE) |>
  select(cdm_name, group_level, var_display, display_value) |>
  pivot_wider(names_from = cdm_name, values_from = display_value)

```

## Baseline Characteristics

### Table 1: Demographics and Clinical Characteristics

```{r baseline-table, echo=FALSE, warning=FALSE, message=FALSE}
baseline_summary |>
  gt(groupname_col = "group_level") |>
  tab_header(title = "Table 1: Baseline Demographics and Clinical Characteristics") |>
  cols_label(var_display = "Characteristic") |>
  tab_style(style = cell_borders(sides = "all", color = "#D3D3D3", weight = px(1)), 
            locations = cells_body()) |>
  tab_style(style = cell_fill(color = "#F0F0F0"), locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#E8E8E8"), locations = cells_row_groups()) |>
  cols_align(align = "left", columns = "var_display") |>
  cols_align(align = "center", columns = -var_display) |>
  tab_options(table.font.size = px(10), heading.align = "left", row_group.font.weight = "bold")
```

### Table 2: Pre-Index Comorbidities (Top 10)

```{r comorbidities-table, echo=FALSE, warning=FALSE, message=FALSE}
comorbidities_pre |>
  mutate(`Target Cohort` = format_cohort_labels(`Target Cohort`)) |>
  gt(groupname_col = "Target Cohort") |>
  tab_header(title = "Table 2: Most Common Comorbidities Prior to Index Date",
             subtitle = "Top 10 conditions per cohort. Format: Database1 | Database2") |>
  tab_style(style = cell_borders(sides = "all", color = "#D3D3D3", weight = px(1)), 
            locations = cells_body()) |>
  tab_style(style = cell_fill(color = "#F0F0F0"), locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#E8E8E8"), locations = cells_row_groups()) |>
  cols_align(align = "left", columns = "Condition") |>
  cols_align(align = "center", columns = "N (%) by Database") |>
  tab_options(table.font.size = px(10), heading.align = "left", row_group.font.weight = "bold")
```

### Table 3: Index Date Admission Types

```{r admissions-table, echo=FALSE, warning=FALSE, message=FALSE}
admissions_index |>
  mutate(`Target Cohort` = format_cohort_labels(`Target Cohort`)) |>
  gt(groupname_col = "Target Cohort") |>
  tab_header(title = "Table 3: Admission Types on Index Date") |>
  tab_style(style = cell_borders(sides = "all", color = "#D3D3D3", weight = px(1)), 
            locations = cells_body()) |>
  tab_style(style = cell_fill(color = "#F0F0F0"), locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#E8E8E8"), locations = cells_row_groups()) |>
  cols_align(align = "left", columns = "Admission Type") |>
  cols_align(align = "center", columns = -c("Target Cohort", "Admission Type")) |>
  tab_options(table.font.size = px(10), heading.align = "left", row_group.font.weight = "bold")
```

### Table 4: Index Date Procedures (Top 10)

```{r procedures-table, echo=FALSE, warning=FALSE, message=FALSE}
procedures_index |>
  mutate(`Target Cohort` = format_cohort_labels(`Target Cohort`)) |>
  gt(groupname_col = "Target Cohort") |>
  tab_header(title = "Table 4: Most Common Procedures on Index Date",
             subtitle = "Top 10 procedures per cohort. Format: Database1 | Database2") |>
  tab_style(style = cell_borders(sides = "all", color = "#D3D3D3", weight = px(1)), 
            locations = cells_body()) |>
  tab_style(style = cell_fill(color = "#F0F0F0"), locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#E8E8E8"), locations = cells_row_groups()) |>
  cols_align(align = "left", columns = "Procedure") |>
  cols_align(align = "center", columns = "N (%) by Database") |>
  tab_options(table.font.size = px(10), heading.align = "left", row_group.font.weight = "bold")
```

### Table 5: Index Date Medications (Top 10)

```{r medications-table, echo=FALSE, warning=FALSE, message=FALSE}
medications_index |>
  mutate(`Target Cohort` = format_cohort_labels(`Target Cohort`)) |>
  gt(groupname_col = "Target Cohort") |>
  tab_header(title = "Table 5: Most Common Medications on Index Date",
             subtitle = "Top 10 medications per cohort. Format: Database1 | Database2") |>
  tab_style(style = cell_borders(sides = "all", color = "#D3D3D3", weight = px(1)), 
            locations = cells_body()) |>
  tab_style(style = cell_fill(color = "#F0F0F0"), locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#E8E8E8"), locations = cells_row_groups()) |>
  cols_align(align = "left", columns = "Medication") |>
  cols_align(align = "center", columns = "N (%) by Database") |>
  tab_options(table.font.size = px(10), heading.align = "left", row_group.font.weight = "bold")
```

### Table 6: Follow-up Period Events (Top 10)

```{r followup-table, echo=FALSE, warning=FALSE, message=FALSE}
bind_rows(followup_conditions, followup_procedures) |>
  select(`Target Cohort` = group_level, `Event Type` = event_type, 
         Event = variable_name, `N (%) by Database` = formatted) |>
  mutate(`Target Cohort` = format_cohort_labels(`Target Cohort`)) |>
  gt(groupname_col = "Target Cohort") |>
  tab_header(title = "Table 6: Most Common Events During Follow-up Period",
             subtitle = "Top 10 conditions and procedures per cohort. Format: Database1 | Database2") |>
  tab_style(style = cell_borders(sides = "all", color = "#D3D3D3", weight = px(1)), 
            locations = cells_body()) |>
  tab_style(style = cell_fill(color = "#F0F0F0"), locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#E8E8E8"), locations = cells_row_groups()) |>
  cols_align(align = "left", columns = c("Event Type", "Event")) |>
  cols_align(align = "center", columns = "N (%) by Database") |>
  tab_options(table.font.size = px(10), heading.align = "left", row_group.font.weight = "bold")
```

```

## Survival Analysis Results
### Summary Table

```{r survival-table, echo=FALSE, warning=FALSE, message=FALSE}
# Build gt table with conditional column alignment
table_col_names <- names(survival_table)
center_cols <- intersect(c("Number records", "Number events", "Median survival (95% CI)", 
                           "Restricted mean survival (95% CI)"), table_col_names)
left_cols <- intersect(c("CDM name", "Target cohort", "Outcome name"), table_col_names)

gt_table <- survival_table |> gt() |>
  tab_header(title = "Survival Analysis Results") |>
  tab_style(style = cell_borders(sides = "all", color = "#D3D3D3", weight = px(1)), locations = cells_body()) |>
  tab_style(style = cell_fill(color = "#F0F0F0"), locations = cells_column_labels()) |>
  tab_options(table.font.size = px(12), heading.align = "left")

if(length(center_cols) > 0) gt_table <- gt_table |> cols_align(align = "center", columns = all_of(center_cols))
if(length(left_cols) > 0) gt_table <- gt_table |> cols_align(align = "left", columns = all_of(left_cols))

gt_table
```

### Survival Curves

```{r survival-plots, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
survival_curves |>
  ggplot(aes(x = time, y = survival_prob, color = target_cohort, fill = target_cohort, linetype = outcome)) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.3, color = NA) +
  geom_line(linewidth = 0.6) +
  facet_wrap(~ cdm_name, scales = "free_x") +
  scale_color_manual(values = c("congenital_aortic_stenosis" = "#F8766D", "congenital_aortic_valve_disease" = "#00BFC4"),
                     name = "Target cohort") +
  scale_fill_manual(values = c("congenital_aortic_stenosis" = "#F8766D", "congenital_aortic_valve_disease" = "#00BFC4"),
                    name = "Target cohort") +
  scale_linetype_manual(values = c("avr" = "solid", "death_cohort" = "dashed"),
                       labels = c("avr" = "AVR", "death_cohort" = "Death"), name = "Outcome") +
  labs(x = "Time in days", y = "Survival probability", title = "Survival Analysis by CDM and Target Cohort") +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "#E0E0E0", color = NA), strip.text = element_text(size = 12, face = "bold"),
        plot.title = element_text(hjust = 0.5),
        legend.position = "right", panel.grid.minor = element_blank(),
        axis.title = element_text(size = 11), axis.text = element_text(size = 10))
```

## Survival Analysis by Age Group

```{r survival-by-age, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
if(nrow(survival_by_age) > 0) {
  survival_by_age |>
    mutate(cohort_label = format_cohort_labels(target_cohort)) |>
    ggplot(aes(x = time, y = survival_prob, color = age_group, fill = age_group, linetype = outcome)) +
    geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.2, color = NA) +
    geom_line(linewidth = 0.6) +
    facet_grid(cohort_label ~ cdm_name, scales = "free", switch = "y") +
    scale_linetype_manual(values = c("avr" = "solid", "death_cohort" = "dashed"),
                         labels = c("avr" = "AVR", "death_cohort" = "Death"), name = "Outcome") +
    labs(x = "Time in days", y = "Survival probability", title = "Survival Analysis by Age Group",
         color = "Age Group", fill = "Age Group") +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "#E0E0E0", color = NA),
          strip.text = element_text(size = 11, face = "bold"),
          strip.placement = "outside",
          plot.title = element_text(hjust = 0.5),
          legend.position = "right", panel.grid.minor = element_blank(),
          axis.title = element_text(size = 11), axis.text = element_text(size = 9))
} else cat("No age-stratified survival data available.")
```

## Survival Analysis by Sex

```{r survival-by-sex, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
if(nrow(survival_by_sex) > 0) {
  survival_by_sex |>
    mutate(cohort_label = format_cohort_labels(target_cohort)) |>
    ggplot(aes(x = time, y = survival_prob, color = sex, fill = sex, linetype = outcome)) +
    geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.2, color = NA) +
    geom_line(linewidth = 0.6) +
    facet_grid(cohort_label ~ cdm_name, scales = "free", switch = "y") +
    scale_color_manual(values = c("Female" = "#FF6B9D", "Male" = "#4A90E2"), name = "Sex") +
    scale_fill_manual(values = c("Female" = "#FF6B9D", "Male" = "#4A90E2"), name = "Sex") +
    scale_linetype_manual(values = c("avr" = "solid", "death_cohort" = "dashed"),
                         labels = c("avr" = "AVR", "death_cohort" = "Death"), name = "Outcome") +
    labs(x = "Time in days", y = "Survival probability", title = "Survival Analysis by Sex") +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "#E0E0E0", color = NA),
          strip.text = element_text(size = 11, face = "bold"),
          strip.placement = "outside",
          plot.title = element_text(hjust = 0.5),
          legend.position = "right", panel.grid.minor = element_blank(),
          axis.title = element_text(size = 11), axis.text = element_text(size = 10))
} else cat("No sex-stratified survival data available.")
```


