@ -0,0 +1,32 @@
---
title: "Temporal Trends in the Real-World Use of SAVR and TAVI in the UK Clinical Practice: A Population-Base Cohort Study from 2012 to 2025"
author:
  - name: Marta Alcalde Herraiz
format:
  docx:
    number-sections: true
    colorlinks: true
    keeptex: true

date: 'last-modified'
date-format: '[This version:] March, 2026'
csl: american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
editor:
  mode: source
filters:
  - docx-landscape.lua
---

```{r}
library(dplyr)
library(omopgenerics)
library(here)
library(CodelistGenerator)
library(readr)
library(visOmopResults)
library(stringr)
library(CohortCharacteristics)
library(tidyr)
library(ggplot2)
library(patchwork)
library(IncidencePrevalence)
source(here("Report", "functions.R"))

result <- importSummarisedResult(path = here("Results")) |>
  mutate("cdm_name" = gsub("HERON_CDM_202509", "CPRD Aurum", cdm_name)) 
codelists_indications <- read_csv(here("Report", "codelists_indications.csv"))
codelists_avr <- read_csv(here("Report", "codelists_avr.csv"))
```
\newpage

```{r}
codelists_indications |>
  inner_join(
    result |>
      filterSettings(result_type == "cohort_code_use") |>
      filterGroup(cohort_name == codelist_name) |>
      splitAll() |>
      select(-"domain_id") ,
    by = c("codelist_name")) |>
  select("cohort_name" = "codelist_name", "concept_id", "concept_name", 
         "domain" = "domain_id", "vocabulary" = "vocabulary_id") |>
  distinct() |>
  mutate("cohort_name" = str_to_sentence(gsub("_", " ", gsub("_avr","",cohort_name)))) |>
  mutate("cohort_name" = factor(cohort_name, levels = c("Aortic stenosis", "Aortic insufficiency", "Aortic endocarditis"))) |>
  arrange(cohort_name, concept_name) |>
  visTable(type = "flextable", 
           groupColumn = "cohort_name", 
           rename = c("Concept ID" = "concept_id"))
```

\newpage
```{r}
codelists_avr |>
  inner_join(
    result |>
      filterSettings(result_type == "cohort_code_use") |>
      filterGroup(cohort_name == codelist_name) |>
      splitAll() |>
      select(-c("domain_id", "codelist_name")) ,
    by = c("cohort_name")) |>
  select("codelist_name", "concept_id", "concept_name", 
         "domain" = "domain_id", "vocabulary" = "vocabulary_id") |>
  distinct() |>
  mutate("codelist_name" = str_to_sentence(gsub("_", " ", gsub("_avr","",codelist_name)))) |>
  mutate("codelist_name" = case_when(
    codelist_name == "Aortic valve replacement potential tavi" ~ "Aortic valve replacement (potential TAVI)",
    codelist_name == "Tavi" ~ "TAVI",
    codelist_name == "Tavi additional" ~ "TAVI (used in combination with an aortic valve replacement -potential TAVI- code)",
    .default = codelist_name
  )) |>
  arrange(codelist_name, concept_name) |>
  visTable(type = "flextable", 
           groupColumn = "codelist_name", 
           rename = c("Concept ID" = "concept_id"))
```




\newpage
Indications by age group (0 to 64)
```{r, fig.width=8, fig.height=12}
p1 <- getStackedPlot(result, 
               cohort_name = "aortic_valve_replacement",
               age_group = "0 to 64", 
               sex = "overall", 
               title = "A) Aortic valve replacement")

p2 <- getStackedPlot(result, 
               cohort_name = "tavi",
               age_group = "0 to 64", 
               sex = "overall", 
               title = "B) Transcatheter aortic valve replacement (TAVI)")

p3 <- getStackedPlot(result, 
               cohort_name = "savr",
               age_group = "0 to 64", 
               sex = "overall", 
               title = "C) Surgical aortic valve replacement (SAVR)")

(p1 / p2 / p3) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

\newpage
Indications by age group (65 to 150)
```{r, fig.width=8, fig.height=12}
p1 <- getStackedPlot(result, 
               cohort_name = "aortic_valve_replacement",
               age_group = "65 to 150", 
               sex = "overall", 
               title = "A) Aortic valve replacement")

p2 <- getStackedPlot(result, 
               cohort_name = "tavi",
               age_group = "65 to 150", 
               sex = "overall", 
               title = "B) Transcatheter aortic valve replacement (TAVI)")

p3 <- getStackedPlot(result, 
               cohort_name = "savr",
               age_group = "65 to 150", 
               sex = "overall", 
               title = "C) Surgical aortic valve replacement (SAVR)")

(p1 / p2 / p3) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

\newpage
Indications by sex (Females)
```{r, fig.width=8, fig.height=12}
p1 <- getStackedPlot(result, 
               cohort_name = "aortic_valve_replacement",
               age_group = "overall", 
               sex = "Female", 
               title = "A) Aortic valve replacement")

p2 <- getStackedPlot(result, 
               cohort_name = "tavi",
               age_group = "overall", 
               sex = "Female", 
               title = "B) Transcatheter aortic valve replacement (TAVI)")

p3 <- getStackedPlot(result, 
               cohort_name = "savr",
               age_group = "overall", 
               sex = "Female", 
               title = "C) Surgical aortic valve replacement (SAVR)")

(p1 / p2 / p3) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

\newpage
Indications by sex (Males)
```{r, fig.width=8, fig.height=12}
p1 <- getStackedPlot(result, 
               cohort_name = "aortic_valve_replacement",
               age_group = "overall", 
               sex = "Male", 
               title = "A) Aortic valve replacement")

p2 <- getStackedPlot(result, 
               cohort_name = "tavi",
               age_group = "overall", 
               sex = "Male", 
               title = "B) Transcatheter aortic valve replacement (TAVI)")

p3 <- getStackedPlot(result, 
               cohort_name = "savr",
               age_group = "overall", 
               sex = "Male", 
               title = "C) Surgical aortic valve replacement (SAVR)")

(p1 / p2 / p3) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

\newpage
Incidence stratified by sex (Female)
```{r, fig.width=8, fig.height=12}
# Incidence
# p1 <- getIncidencePlot(result, age_group = c("40 to 64", "65 to 69"), sex = "Female")
# p2 <- getIncidencePlot(result, age_group = c("70 to 74", "75 to 79"), sex = "Female")
# p3 <- getIncidencePlot(result, age_group = c("80 to 84", "85 to 150"), sex = "Female")
# 
# p <- (p1/p2/p3) + 
#   plot_layout(guides = "collect") &
#   theme(legend.position = "top") 
```

\newpage
Incidence stratified by sex (Male)
```{r, fig.width=8, fig.height=12}
# Incidence
# p1 <- getIncidencePlot(result, age_group = c("40 to 64", "65 to 69"), sex = "Male")
# p2 <- getIncidencePlot(result, age_group = c("70 to 74", "75 to 79"), sex = "Male")
# p3 <- getIncidencePlot(result, age_group = c("80 to 84", "85 to 150"), sex = "Male")
# 
# p <- (p1/p2/p3) + 
#   plot_layout(guides = "collect") &
#   theme(legend.position = "top") 
```

\newpage
Table incidence (Aortic valve replacement) - Overall
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "aortic_valve_replacement", 
                  age_group = c("0 to 150"), 
                  sex = "Both") 
```

\newpage
Table incidence (TAVI) - Overall
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "tavi", 
                  age_group = c("0 to 150"), 
                  sex = "Both") 
```

\newpage
Table incidence (SAVR) - Overall
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "savr", 
                  age_group = c("0 to 150"), 
                  sex = "Both") 
```

\newpage
Table incidence (Aortic valve replacement) - By age
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "aortic_valve_replacement", 
                  age_group = c("0 to 39", "40 to 64", "65 to 69", "70 to 74", "75 to 79", "80 to 84", "85 to 150"), 
                  sex = "Both") 
```

\newpage
Table incidence (Tavi) - By age
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "tavi", 
                  age_group = c("0 to 39", "40 to 64", "65 to 69", "70 to 74", "75 to 79", "80 to 84", "85 to 150"), 
                  sex = "Both") 
```

\newpage
Table incidence (savr) - By age
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "savr", 
                  age_group = c("0 to 39", "40 to 64", "65 to 69", "70 to 74", "75 to 79", "80 to 84", "85 to 150"), 
                  sex = "Both") 
```

\newpage
Table incidence (aortic valve replacement) - By sex 
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "aortic_valve_replacement", 
                  age_group = c("0 to 150"),
                  sex = c("Female", "Male")) 
```

\newpage
Table incidence (tavi) - By sex 
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "tavi", 
                  age_group = c("0 to 150"),
                  sex = c("Female", "Male")) 
```

\newpage
Table incidence (savi) - By sex 
```{r}
getIncidenceTable(result,
                  outcome_cohort_name = "savr", 
                  age_group = c("0 to 150"),
                  sex = c("Female", "Male")) 
```

\newpage
Table incidence (aortic valve replacement) - By age and sex 
```{r}
# getIncidenceTable(result,
#                   outcome_cohort_name = "aortic_valve_replacement", 
#                   age_group = c("0 to 39", "40 to 64", "65 to 69", "70 to 74", "75 to 79", "80 to 84", "85 to 150"), 
#                   sex = c("Female", "Male")) 
```

\newpage
Table incidence (tavi) - By age and sex 
```{r}
# getIncidenceTable(result,
#                   outcome_cohort_name = "tavi", 
#                   age_group = c("0 to 39", "40 to 64", "65 to 69", "70 to 74", "75 to 79", "80 to 84", "85 to 150"), 
#                   sex = c("Female", "Male")) 
```

\newpage
Table incidence (savr) - By age and sex 
```{r}
# getIncidenceTable(result,
#                   outcome_cohort_name = "savr", 
#                   age_group = c("0 to 39", "40 to 64", "65 to 69", "70 to 74", "75 to 79", "80 to 84", "85 to 150"), 
#                   sex = c("Female", "Male")) 
```

